### 2024-03-15 article.Rmd
revisions to article

---
title: "sfislands: An R Package for Accommodating Islands and Disjoint Zones in Areal Spatial Modelling"
date: "2022-11-09"
abstract: >
  Fitting areal spatial models can be a cumbersome task, particularly when the geographical units are not well-behaved. The presence of islands gives rise to particular issues when creating neighbourhood structures based on contiguity. We might also want to edit the output of a condition-based neighbourhood structure. It might be appropriate to add or remove connections to take account of infrastructure links or other domain-specific issues which are not apparent from a basic set of polygons. This package streamlines these processes with tools to account for the presence of islands and to intuitively make manual alterations to neighbourhood structures with the help of convenient mapping functions. Beyond the setting up of a satisfactory neighbourhood structure, it has further helper functions specific to the `mgcv` package which reduce the human workload required to extract estimates from models with spatial and hierarchical structures.
draft: true
author:  
  - name: Kevin Horan
    affiliation: 
    - Hamilton Institute, Maynooth University
    address:
    - Maynooth
    - Co. Kildare, Ireland
    url: https://github.com/horankev
    orcid: 0009-0003-9378-0084
    email:  kevin.horan.2021@mumail.ie
  - name: Katarina Domijan
    affiliation:
    - Department of Mathematics and Statistics, Maynooth University
    address:
    - Maynooth
    - Co. Kildare, Ireland
    email: katarina.domijan@mu.ie
    orcid: 0000-0002-4268-2236
  - name: Chris Brunsdon
    affiliation:
    - National Centre for Geocomputation, Maynooth University
    address:
    - Maynooth
    - Co. Kildare, Ireland
    email: Christopher.Brunsdon@mu.ie
    orcid: 0000-0003-4254-1780
type: package
output: 
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center")

# install.packages("devtools")
devtools::install_github("horankev/sfislands")

# required libraries
library(tidyverse)
library(kableExtra)
library(sf)
library(sfdep)
library(sfislands)
library(mgcv)
library(here)
library(ggsflabel)
library(ggspatial)
library(ggpubr)
library(AER) # to test for overdispersion
library(ggmagnify)

# To scrape data
library(rvest)
library(httr)
library(polite)
library(janitor)

# for geom_sf_label_repel consistency
set.seed(123)

# set consistent theme
theme_set(
  theme_bw() +
    theme(panel.background = element_rect(fill = "#F6F3E9", colour = "black", linewidth=1.5), 
          axis.title = element_blank(),
          axis.ticks = element_blank())
)

```

```{r}
# load data

# datasets are all available at https://github.com/horankev/quake_data 
# or https://github.com/horankev/london_liverpool_data
# alongside details on precisely how they were sourced etc

# indonesia
quakes_df <- readRDS(gzcon(url("https://github.com/horankev/quake_data/raw/main/datasets/quakes_df.rds")))
provinces_df <- readRDS(gzcon(url("https://github.com/horankev/quake_data/raw/main/datasets/provinces_df.rds"))) |> 
  rename(damaging_quakes_total = quake_mlxl_total,
         damaging_quakes_density = quake_mlxl_density)
counties_df <- readRDS(gzcon(url("https://github.com/horankev/quake_data/raw/main/datasets/counties_df.rds")))
faults_df <- readRDS(gzcon(url("https://github.com/horankev/quake_data/raw/main/datasets/faults_df.rds")))
nearby_countries_df <- readRDS(gzcon(url("https://github.com/horankev/quake_data/raw/main/datasets/nearby_countries_df.rds")))
indonesia_outline <- provinces_df |> summarise()

# london
london <- readRDS(gzcon(url("https://github.com/horankev/london_liverpool_data/raw/main/data/london.rds")))
thames <- readRDS(gzcon(url("https://github.com/horankev/london_liverpool_data/raw/main/data/thames.rds")))
crossings_roadped <- readRDS(gzcon(url("https://github.com/horankev/london_liverpool_data/raw/main/data/crossings_roadped.rds")))

# liverpool
liverpool <- readRDS(gzcon(url("https://github.com/horankev/london_liverpool_data/raw/main/data/liverpool.rds")))
msoa_sf <- liverpool |> 
  group_by(msoa_cd) |> 
  summarise()
lsoa_sf <- liverpool |> 
  group_by(lsoa_cd) |> 
  summarise()
mersey <- readRDS(gzcon(url("https://github.com/horankev/london_liverpool_data/raw/main/data/mersey.rds")))

```

# Introduction

A key feature which differentiates spatial statistics is the
non-independence of observations and the expectation that neighbouring
units will be more similar than non-neighbouring ones [@tobler]. If this
is not accounted for, the assumptions of many types of models will be
violated. The relationships between all spatial units in a study can be
represented numerically in a spatial weights matrix. In order to build
this, we must first decide on what constitutes being a neighbour. We can
see this as a continuous relationship where degree of neighbourliness is
a function of connectivity, which could be represented as some measure
of distance. Alternatively it could be a binary situation where each
pair of units either are (1) or are not (0) neighbours. This can be
based on a condition such as contiguity of some sort, or a distance
constraint. It is the job of the modeller to formulate a hypothesis
which justifies their choice of neighbourhood structure.

For R users, the \CRANpkg{spdep} package [@spdep] has long been popular for the
creation of these matrices. More recently, with the increasing use of \CRANpkg{sf}
structures and their compatibility with the \CRANpkg{tidyverse}  [@tidyverse], the \CRANpkg{sfdep} package [@sfdep] has presented the same
functionality in addition to extra features in a tidy structure based on
lists in a similar way to \CRANpkg{sf}.

The most appropriate form of neighbourhood structure will depend on the specific context. @BrizRedn2021 compared different structures in the context of COVID-19 data. They note that @Earnest2007 found that distance-based matrices were more appropriate when examining birth defects in Australia, whereas @Duncan2017 found that a first-order contiguity structure produced a better fit than others in the context of lip cancer incidence in Scotland.

The most commonly used neighbourhood structure is one based on first-order queen contiguity, where units are considered neighbours if they
share at least a vortex of boundary. However, as the name suggests, this
will lead to problems when non-contiguous units such as islands or
exclaves are present. Less obviously, depending on how the geographic
units are described, areas on either sides of rivers may be
inappropriately classified as neighbours or not neighbours. Furthermore,
the presence of infrastructure such as tunnels, bridges or ferry
services might be satisfactory to meet our hypothesis of the required
degree of connectivity to be considered neighbours. Again, such information will not be found in a set of polygons.

The aim of `sfislands` is to deal with the situations described above in a
convenient and open manner. It allows us to set up a structure, quickly
map it, and then examine whether or not it is consistent with our hypothesis of relationships between units. The structure can then be edited and the process re-iterated until we have described a
spatial relationship structure with which we are satisfied.

The above can be considered as the *pre-functions* of the package. A
second category of features, which we refer to as *post-functions*,
are for use with the \CRANpkg{mgcv} package [@mgcv]. Having fit a model with \CRANpkg{mgcv}, the process of extracting estimates for certain types of effects can be somewhat awkward. These *post-functions* augment the original dataframe with these estimates and their standard errors in tidy format. They also allow for quick visualisation of the output in map form.

## Typical use-cases

In this paper, we will look at three examples to show different use-cases for
`sfislands`. The first example focuses on earthquakes in
Indonesia. It shows a scenario where all of the functions are used, from
setting up contiguities, to modelling and examining the estimates of the
model.

The second example looks at London and how, despite the lack of islands,
the presence of a river means that some of the pre-functions of
`sfislands` can be useful.

The final example focuses on Liverpool. There are no islands or issues
of discontiguity in this dataset. Instead, we will show how the package
can be used to fit a multilevel model with an ICAR component at the lowest
level. We show how the results can be presented in tidy form and quickly
visualised. We also show how different types of contiguity from the
\CRANpkg{sfdep} package can used within the \CRANpkg{mgcv} modelling
framework and how the estimates derived from them can be conveniently
compared.

# Why use `sfislands`?

It greatly speeds up each stage of a proposed workflow for fitting areal spatial models.

### Step 1: *Pre-functions* for setting up neighbourhood structure

  1. It addresses an issue commonly seen in online help forums where an inexperienced user wishes to get started with a model but fails at the first hurdle because their neighbourhood structure is not compatible with their data. `sfislands` will include a contiguity for all units.

  2. It gives tools to immediately visualise this structure as a map.

  3. These maps are created using \CRANpkg{ggplot2}  [@ggplot2], which allows users to apply the desired styling and themes using \CRANpkg{ggplot2} syntax.

  4. As the nodes can be labelled by index, it makes it very easy to add and remove connections as appropriate with confidence.

  5. Connections which have been induced by a function from the package but are not contiguities can be accessed to ensure openness in the process.

### Step 2: Modelling

  These neighbourhood structures can be used in modelling packages such as \CRANpkg{mgcv}, \CRANpkg{brms}  [@brms], \pkg{r-inla}  [@r-inla] and more.

### Step 3: *Post-functions* for `mgcv` models

  1. It simplifies the process of extracting estimates from models with random effects and Markov random field structures (ICAR models).

  2. These effects can be quickly visualised as \CRANpkg{ggplot2} maps.

# Pre-functions

The first group of functions, shown in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:prefunc-html)', '\\@ref(tab:prefunc-latex)'))`, deals with the creation
of a neighbourhood structure in the presence of discontiguities. The
resultant structure can be quickly mapped to check if it is
satisfactory. Connections can be manually added or removed by name or
index number. By an iterative process of changes and examination of a
quickly-generated guide map, a satisfactory structure can be decided upon.

```{r}
prefunc <- tribble(
  ~`function`, ~purpose,
  "st_bridges()", "create a neighbourhood contiguity structure, with a k-nearest neighbours condition for islands",
  "st_quickmap_nb()", "check structure visually on map",
  "st_check_islands()", "check the contiguities which have been assigned to islands",
  "st_manual_join_nb()", "make manual changes by adding connections",
  "st_manual_cut_nb()", "make manual changes by removing connections"
)

```

```{r prefunc-html, eval = knitr::is_html_output(), layout = "center"}
knitr::kable(prefunc, format = "html", caption = "Pre-functions: setting up a neighbourhood structure.") |> 
  kable_styling(full_width = TRUE)
```

```{r prefunc-latex, eval = knitr::is_latex_output()}
knitr::kable(prefunc, format = "latex", caption = "Pre-functions: setting up a neighbourhood structure.") |> 
  row_spec(0, bold = TRUE) |> 
  kableExtra::kable_styling(font_size = 9) |> 
  column_spec(2, width = "7cm")
```

We will now go through each function in more detail using the set of rectangles shown in Figure \@ref(fig:rects1) for demonstration purposes. Rectangles 1, 2 and 3 are contiguous while 4 and 5 can be viewed as "islands".

```{r}
rect1 <- st_polygon(list(rbind(c(0, 0), c(0, 2), c(2, 2), c(2, 0), c(0, 0))))
rect2 <- st_polygon(list(rbind(c(2, 0), c(2, 2), c(4, 2), c(4, 0), c(2, 0))))
rect3 <- st_polygon(list(rbind(c(2, 2), c(2, 4), c(4, 4), c(4, 2), c(2, 2))))
rect4 <- st_polygon(list(rbind(c(5, 0), c(5, 1), c(6, 1), c(6, 0), c(5, 0))))
rect5 <- st_polygon(list(rbind(c(0.8, 3), c(0.8, 4), c(1.8, 4), c(1.8, 3), c(0.8, 3))))

# Create sf objects for each rectangle
rect1_sf <- st_sf(geometry = st_sfc(rect1))
rect2_sf <- st_sf(geometry = st_sfc(rect2))
rect3_sf <- st_sf(geometry = st_sfc(rect3))
rect4_sf <- st_sf(geometry = st_sfc(rect4))
rect5_sf <- st_sf(geometry = st_sfc(rect5))

# Combine the rectangles into one sf object
rectangles <- rbind(rect1_sf, rect2_sf, rect3_sf, rect4_sf, rect5_sf) |> 
  mutate(name = c("Rect1","Rect2","Rect3","Rect4","Rect5"))
```

```{r, rects1, fig.width=4, fig.height=2, fig.cap = "Simplified scenario with five rectangles. "}
ggplot(rectangles) + geom_sf(colour="black", linewidth=1) + geom_sf_label(aes(label=name),size=2.5) + 
  coord_sf(datum=NA) + 
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_rect(fill = "#ECF6F7", colour = "black", linewidth=1.5))

```

## st_bridges()

This function requires at least two arguments: an \CRANpkg{sf} dataframe and, from that, the name of one column of unique identifiers, ideally names, of each spatial unit. It creates a neighbourhood structure where non-island units are joined by first-order queen contiguity, while island units are joined to their k-nearest neighbours. The output is a *named* neighbourhood structure in either list or matrix form as desired, which can be either a standalone object or included as an additional column in the original \CRANpkg{sf} dataframe. While it is not necessary in all modelling packages for the neighbourhood list or matrix to be *named*, it is good practice to do so and is mandatory when using, for example, \CRANpkg{mgcv}.

In the first two examples of using `st_bridges()`, we have chosen to ignore islands with the argument `remove_islands = TRUE` and to return a list and matrix structure respectively by specifying this in the `nb_structure` argument and choosing `add_to_dataframe = FALSE`:

```{r, echo=TRUE}
# output a named list

st_bridges(rectangles, 
           "name", 
           remove_islands = TRUE, 
           nb_structure = "list", 
           add_to_dataframe = FALSE) |>
  head()

```

```{r, echo=TRUE}
# output a named matrix

st_bridges(rectangles, 
           "name", 
           remove_islands = TRUE, 
           nb_structure = "matrix", 
           add_to_dataframe = FALSE) |> 
  head()

```

Alternatively, in the following examples, we choose to join islands to their 1 nearest
neighbour, which is the default setting, and to return the output as a column called "nb" in the original
\CRANpkg{sf} dataframe (`add_to_dataframe = "TRUE"` is the default setting):

```{r, echo=TRUE}
# output a named list as a column "nb" in original dataframe

st_bridges(rectangles, 
           "name", 
           link_islands_k = 1, 
           nb_structure = "list") |> 
  head()

```

```{r, echo=TRUE}
# output a named matrix as a column "nb" in original dataframe

st_bridges(rectangles, 
           "name", 
           link_islands_k = 1, 
           nb_structure = "matrix") |> 
  head()

```

These structures can serve as the input to models in \CRANpkg{brms},
\pkg{r-inla}, \pkg{rstan}  [@rstan] or \CRANpkg{mgcv}. \CRANpkg{brms} requires a matrix structure while \CRANpkg{mgcv} models use a list. Rather than having a separate neighbours
object, it is included in the original \CRANpkg{sf} dataframe as a named list or matrix, in the spirit of the \CRANpkg{sfdep} package.

## st_quickmap_nb()

It is much more intuitive to examine these structures visually than in
matrix or list format. This can be done with the `st_quickmap_nb()`
function as shown in Figure \@ref(fig:rects2).

```{r, rects2, fig.width=2, fig.height=1, echo=TRUE, fig.cap = "Queen contiguity and islands connected to nearest neighbour. "}
# default is 'nodes = "point"'

st_bridges(rectangles, 
           "name", 
           link_islands_k = 1) |> 
  st_quickmap_nb()

```

If we wish to make edits, it might be more useful to represent the nodes
numerically rather than as points (Figure \@ref(fig:rects3)).

```{r, rects3, fig.width=2, fig.height=1, echo=TRUE, fig.cap = "Queen contiguity and islands connected to nearest neighbour. Nodes are shown as numeric indices. "}
# with 'nodes = "numeric"'

st_bridges(rectangles, 
           "name", 
           link_islands_k = 1) |> 
  st_quickmap_nb(nodes = "numeric")

```

## st_check_islands()

This function will show us transparently what connections have been made which
are not based on contiguity. It gives both the name and index number of each pair of added connections. In this example, two pairs have been added.

```{r, echo=TRUE}
# show summary of non-contiguous connections in a dataframe

st_bridges(rectangles, 
           "name", 
           link_islands_k = 1) |> 
  st_check_islands()

```

## st_manual_join_nb()

If we feel that 4 should also be connected to 3, this can be done
manually (Figure \@ref(fig:rects4)).

```{r, rects4, fig.width=2, fig.height=1, echo=TRUE, fig.cap = "With an additional connection between 3 and 4. "}
# add an extra connection using numeric index

st_bridges(rectangles, "name", 
           link_islands_k = 1) |> 
  st_manual_join_nb(3,4) |> 
  st_quickmap_nb(nodes = "numeric")
```

## st_manual_cut_nb()

And perhaps there is a wide river between rectangles 1 and 2 which
justifies removing the connection. We will edit it this time using
names (Figure \@ref(fig:rects5)).

```{r, rects5, fig.width=2, fig.height=1, echo=TRUE, fig.cap = "With an additional connection between 1 and 2. "}
# remove an existing connection using unit name, not index

st_bridges(rectangles, "name", 
           link_islands_k = 1) |> 
  st_manual_join_nb(3,4) |> 
  st_manual_cut_nb("Rect1","Rect2") |> 
  st_quickmap_nb(nodes = "numeric")
```

The use of these neighbourhood structures is particularly associated with CAR-type models.
These are often implemented in a Bayesian framework using
\CRANpkg{brms}, \pkg{r-inla} or \pkg{rstan}. For example, the \CRANpkg{brms} ICAR structure requires the neighbourhood relationships to be in matrix form. The pre-functions will output the
neighbourhood structure in the desired format for use in any of these
frameworks. A convenient frequentist alternative is to use the
\CRANpkg{mgcv} package which requires a named list of neighbours. It has the functionality to create such models
using `bs="mrf"`. It also has the ability to combine these with a
hierarchical structure using `bs="re"`. While the outputs from the
Bayesian structures mentioned above can be extracted in the same way as
any other component of the model, it can be somewhat awkward to get the
estimates from \CRANpkg{mgcv} model. `sfislands` has two post-functions to conveniently extract and visualise these.

# Post-functions

Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:postfunc-html)', '\\@ref(tab:postfunc-latex)'))` shows the second set of functions in the package and their purpose.

```{r}
postfunc <- tribble(
  ~`function`, ~purpose,
  "st_augment()", "augment the original dataframe with model predictions",
  "st_quickmap_preds()", "generate quick maps of these predictions"
)

```

```{r postfunc-html, eval = knitr::is_html_output(), layout = "center"}
knitr::kable(postfunc, format = "html", caption = "Post-functions: tidy estimates from `mgcv`.") |> 
  kable_styling(full_width = TRUE)
```

```{r postfunc-latex, eval = knitr::is_latex_output()}
knitr::kable(postfunc, format = "latex", caption = "Post-functions: tidy estimates from mgcv.") |> 
  row_spec(0, bold = TRUE) |> 
  kableExtra::kable_styling(font_size = 9, full_width = FALSE)
```

## st_augment()

This function augments the original dataframe with the estimated means
and standard errors of the spatially varying predictions from a fitted \CRANpkg{mgcv} model in a similar manner to
how the \CRANpkg{broom} package  [@broom] operates. The `geometry` column, as per convention, remains as the last column of the augmented dataframe, while the predictions are positioned immediately before it. ^[In a similar way, `st_augment()` can also be used to append the random effects from `lme4` and `nlme` models to an `sf` dataframe, which can then be easily mapped using `st_quickmap_preds()`.] The spatially varying predictions which `st_augment()` extracts from an \CRANpkg{mgcv} model are

*   random effects (which are called in \CRANpkg{mgcv} with `bs='re'`), and 
*   ICAR components (`bs='mrf'`).

Consider the model structure described in the code below using \CRANpkg{mgcv} syntax. In this model $y$ is the dependent variable while it is being estimated with a fixed intercept, a fixed slope for some covariate, a set of random intercepts and slopes for the covariate at a *region* level, and a set of ICAR varying intercepts and slopes at a lower *sub-region* level.

```{r, eval=FALSE, echo=TRUE}
# creating an mgcv model

mgcv::gam(
  y ~ covariate +                   # fixed intercept and effect for covariate  
    s(region, bs = "re") +                  # random intercept at level region
    s(region, covariate, bs = "re") +          # random slopes at level region
    s(sub-region, 
      bs = 'mrf',
      xt = list(nb = data$nb), 
      k = k) +                    # ICAR varying intercept at level sub-region
    s(sub-region, by = covariate, 
      bs = 'mrf',
      xt = list(nb = data$nb), 
      k = k),           # ICAR varying slope for covariate at level sub-region
  data = data, 
  method = "REML")
```

When labelling the new prediction columns which are augmented to the original dataframe from such a model, `st_augment()` follows the formula syntax of the \CRANpkg{lme4} package  [@lme4], where the pipe symbol (`|`) indicates "*grouped by*". Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:staugtab-html)', '\\@ref(tab:staugtab-latex)'))` shows how the augmented columns in this scenario would be named. Each column name begins with either `random.effect.` or `mrf.smooth.` as appropriate. An additional column is also added for the standard error of each estimate, as calculated by \CRANpkg{mgcv}. These columns are named as above but with `se.` prepended (e.g. `se.random.effect.region`).

```{r}
st_aug_tab_html <- tribble(
  ~`mgcv syntax`, ~`column name`,
  "s(region, bs = 're')", "random.effect.region",
  "s(region, covariate, bs = 're')", "random.effect.covariate|region",
  "s(sub-region, bs = 'mrf', xt = list(nb = data\\$nb)", "mrf.smooth.sub-region",
  "s(sub-region, by = covariate, bs = 'mrf', xt = list(nb = data\\$nb)", "mrf.smooth.covariate|sub-region"
)
```

```{r}
st_aug_tab_latex <- tribble(
  ~`mgcv syntax`, ~`column name`,
  "s(region, bs = 're')", "random.effect.region",
  "s(region, covariate, bs = 're')", "random.effect.covariate|region",
  "s(sub-region, bs = 'mrf', xt = list(nb = data$nb)", "mrf.smooth.sub-region",
  "s(sub-region, by = covariate, bs = 'mrf', xt = list(nb = data$nb)", "mrf.smooth.covariate|sub-region"
)
```

```{r staugtab-html, eval = knitr::is_html_output(), layout = "center"}
knitr::kable(st_aug_tab_html, format = "html", caption = "The naming procedure for augmented columns from different `mgcv` structures.") |> 
  kable_styling(full_width = TRUE)
```

```{r staugtab-latex, eval = knitr::is_latex_output()}
knitr::kable(st_aug_tab_latex, format = "latex", caption = "The naming procedure for augmented columns from different mgcv structures.") |> 
  row_spec(0, bold = TRUE) |> 
  kableExtra::kable_styling(font_size = 9, full_width = FALSE, )
```

## st_quickmap_preds()

These estimates can then be quickly mapped. As it is possible to include more than 1 spatially varying component, the output of this function is a list of plots. They can be viewed individually by indexing, or all at one using, for example, the `plotlist` argument from the `ggarrange()` function which is part of the \CRANpkg{ggpubr}  [@ggpubr] package. We will see this function in practice in the following examples. The maps which it generates are automatically titled and subtitled according to the type of effect. For example, the map showing predictions for `random.effect.region` will have "*region*" as its title and "*random.effect*" as its subtitle.

# Example 1: Indonesia

Modelling earthquakes in Indonesia serves as a good example to
demonstrate this package. Firstly, Indonesia is composed of many
islands. Secondly, earthquake activity is known to be associated with
the presence of faults which exist below sea level and thus do not
respect land boundaries. Therefore it is reasonable to expect similar
behaviour in nearby provinces regardless of whether or not they are
contiguous. We aim to model the incidence, or count per unit area, of earthquake activity by
province across Indonesia, controlling for proximity to faults.

## Data

The data for this section have been downloaded from the National Earthquake Information Center, [USGS earthquake
catalogue](https://earthquake.usgs.gov/earthquakes/search/). The
datasets with accompanying explanations are available at
<https://github.com/horankev/quake_data>. They capture all recorded earthquakes in
and close to Indonesia from the beginning of January 1985 to the end of December 2023. Figure \@ref(fig:fault-buffers) shows a map of Indonesia, divided into 33 provinces, with other neighbouring or bordering countries filled in grey. The many local
faults which lie within 300km of the shore are shown in yellow with green outlines.

```{r, fault-buffers, fig.width=12, fig.height=6, out.width="100%", fig.cap = "Indonesia faults. Surrounded by a 10 kilometre buffer. "}

faults_buf <- faults_df |> 
  st_buffer(dist = 10000)

ggplot() + 
  geom_sf(data=nearby_countries_df, fill="gray50", linewidth=0.5, colour="black") +
  geom_sf(data = provinces_df, fill="antiquewhite", colour="black",linewidth=0.5) +
  geom_sf(data = faults_buf, fill="darkgreen", colour="gray50",linewidth=0.9) +
  geom_sf(data=faults_df, colour="yellow", linewidth=0.4) +
  annotation_scale() +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", linewidth=1.5),
        axis.title = element_blank(),
        axis.ticks = element_blank())

```

To get an interpretable measure of the concentration of faults in any area,
these faults are transformed from linestrings to polygons by setting a buffer of
10km around them, which explains their green outline. Now both our faults and the sizes of provinces are in units of kilometres
squared. This means we can generate a unitless metric of what proportion of any administrative unit
is covered by these buffered faults. This measure across provinces is shown in Figure
\@ref(fig:fault-conc).

```{r, fault-conc, fig.width=12, fig.height=6, out.width="100%", fig.cap = "Indonesia fault concentration. Square kilometre of buffered fault per square kilometre of province area. "}

ggplot() + 
  geom_sf(data=nearby_countries_df, fill="gray50", colour="black", linewidth=0.5) +
  geom_sf(data=provinces_df, aes(fill=fault_concentration), colour="black", linewidth=0.5) +
  geom_sf(data=faults_df, colour="gray10", linetype="dashed", linewidth=0.5) +
  scale_fill_distiller(palette = "YlGn", direction = 1) + 
  labs(fill="fault\nconcentration") + 
  annotation_scale() +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", linewidth=1.5),
        legend.position = c(0.92,0.8),
        legend.box.background = element_rect(colour = "black", linewidth = 1))

```

Earthquake incidence per province has been calculated as the total number of earthquakes with
an epicentre within that province per unit area. We have
restricted counts to earthquakes \>5.5 on the moment magnitude scale, which is the point
at which they are often labelled as potentially damaging.

The occurrences of these earthquakes are shown in Figure \@ref(fig:quake-occur), their total per province in Figure \@ref(fig:quake-totals), and finally, their incidence or count per square kilometre can be seen in Figure \@ref(fig:quake-conc).

```{r, quake-occur, fig.width=12, fig.height=6, out.width="100%", fig.cap = "Earthquakes in Indonesia of magnitude > 5.5, 1985-2023. Categorised by magnitude as medium, large or extra-large. "}

quakes_df_temp <- quakes_df |> 
  mutate(magfact = case_when(magfact == "M" ~ "M: 5.5 - 6.1",
                             magfact == "L" ~ "L: 6.1 - 7",
                             magfact == "XL" ~ "XL: 7+")) |> 
  mutate(magfact = factor(magfact,
                          levels = c("M: 5.5 - 6.1",
                                     "L: 6.1 - 7",
                                     "XL: 7+")))
ggplot() + 
  geom_sf(data=nearby_countries_df, fill="gray50", linewidth=0.5, colour="black") +
  geom_sf(data=provinces_df, fill="antiquewhite", colour="black", linewidth=0.5) + 
  geom_sf(data=quakes_df_temp |> filter(magfact != "S") |> filter(province != "undersea"), 
          aes(fill=magfact, size=magfact), shape=21, colour="black") +
  scale_fill_manual(values = c("white","khaki","tomato4")) +
  scale_size_manual(values = c(3.5,5.5,7.5)) + 
  labs(fill="magnitude",
       size="magnitude") + 
  annotation_scale() +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", linewidth=1.5),
        legend.position = c(0.92,0.8),
        legend.key = element_blank(),
        legend.box.background = element_rect(colour = "black", linewidth = 1))

```

```{r, quake-totals, fig.width=12, fig.height=6, out.width="100%", fig.cap = "Earthquake count in Indonesia, 1985-2023, mag > 5.5: count by province. "}

ggplot() +
  geom_sf(data=nearby_countries_df, fill="gray50", colour="black", linewidth=0.5) +
  geom_sf(data=provinces_df, aes(fill=damaging_quakes_total), colour="black", linewidth=0.5) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) + 
  labs(fill="count") + 
  annotation_scale() +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", linewidth=1.5),
        legend.position = c(0.92,0.78),
        legend.box.background = element_rect(colour = "black", linewidth = 1))

```

```{r, quake-conc, fig.width=12, fig.height=6, out.width="100%", fig.cap = "Earthquake incidence in Indonesia, 1985-2023, mag > 5.5: count per square kilometre by province. "}

  ggplot() +
  geom_sf(data=nearby_countries_df, fill="gray50", colour="black", linewidth=0.5) +
  geom_sf(data=provinces_df, aes(fill=damaging_quakes_density), colour="black", linewidth=0.5) +
  scale_fill_distiller(palette = "YlOrRd", direction = 1) + 
  labs(fill="incidence") + 
  annotation_scale() +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", linewidth=1.5),
  axis.title = element_blank(),
        axis.ticks = element_blank(),
        legend.position = c(0.92,0.78),
        legend.box.background = element_rect(colour = "black", linewidth = 1))

```

## Model

As this is count data, we will model it as a poisson distribution with
$\lambda$ as the mean count per province. For $i = 1,...,n$ provinces,
the dependent variable in this model is
\begin{equation}
y_i = \text{earthquake count}_i 
(\#eq:eq1)
\end{equation}
while the explanatory variable is
\begin{equation}
x_i = \text{fault concentration}_i = \frac{\text{area of buffered faults in province}_i} {\text{province area}_i}.
(\#eq:eq2)
\end{equation}
Firstly, when excluding the incidence and just modelling counts, where
$y_i = \text{earthquake count in province}_i$, the poisson model is of
the following form: 
\begin{equation}
y_i | \lambda_i \sim \text{Pois}(\lambda_i)
(\#eq:eq3)
\end{equation}
with
\begin{equation}
E(y_i | \lambda_i) = \lambda_i.
(\#eq:eq4)
\end{equation}
We model
\begin{equation}
log(\lambda_i) = \beta_0 + \beta_1x_i + \gamma_i.
(\#eq:eq5)
\end{equation}
where $\gamma_i$
is a term with a correlation structure reflecting a province’s location
relative to other provinces.

We can describe these relationships by setting up a neighbourhood
structure based on queen contiguity where a pair of provinces are
considered neighbours if they share at least one point of boundary. This can be modelled as a Markov random field to generate an ICAR model with a spatially varying term. Each of these terms will be correlated with the others according to the neighbourhood structure we have defined.

The Markov random field here follows a multivariate Gaussian
distribution. $\gamma_i$ is a vector of province effects having a
distribution with mean **0** and precision **P** where

$[\textbf{p}]_{ij} = v_i$ if $i=j$ and $v_i$ is the number of adjacent provinces
to province $i$,

$[\textbf{p}]_{ij} = -1$ if provinces $i$ and $j$ are adjacent, and

$[\textbf{p}]_{ij} = 0$ otherwise.

A further constraint that $\Sigma_j \gamma_j = 0$ is applied so that the
distribution is identifiable.

We now include an offset term (here, area) because we are more interested
in modelling the incidence than in the actual count, such that
\begin{equation}
log(\frac{\lambda_i} {\text{area}_i}) = \beta_0 + \beta_1x_i + \gamma_i
(\#eq:eq6)
\end{equation}
which is equivalent to
\begin{equation}
log(\lambda_i) = \beta_0 + \beta_1x_i + \gamma_i + log(\text{area}_i).
(\#eq:eq7)
\end{equation}
We are still modelling $log(\lambda)$ rather than the incidence, but we
are adding an offset to adjust for differing areas. Modelling $log(\lambda)$ and adding an offset is equivalent to
modelling incidence, and coefficients can be interpreted that way.

When interpreting the estimated coefficients of the model, it can be
useful to look at it in the following form:
\begin{equation}
\lambda_i = e^{\beta_0 + \beta_1 x_i + \gamma_i} \text{area}_i.
(\#eq:eq8)
\end{equation}

Having described the type of model we wish to implement, we now show how `sfislands` can be used to streamline the process.

## Pre-functions

Such models, however, can not deal with locations which have no
neighbours. In the case of Indonesia, this is quite problematic. It is
composed of many islands. The estimated count of
islands according to @Andrfout2022 is 13,558.
While it is not unusual for a country to have a number of often small
offshore islands, Indonesia is entirely composed of (at least portions
of) an archipelago of islands, so many of these islands or groups of
islands are individual provinces in their own right. We might like to hypothesise that just because a province is a disconnected island, this should not mean that it is independent of other nearby provinces in terms of earthquake incidence. A standard first-order queen contiguity structure would mean the exclusion of disconnected units entirely from the model. An alternative strategy of assigning neighbour status based on a distance metric would overcome this, but the threshold size of distance necessary for such a structure might be inappropriately large for the non-islands provinces. We would like to use a compromise between these two strategies.

In this case, we use `st_bridges()` for setting up the queen contiguity
structure as usual, but with the additional stipulation that unconnected
units (provinces which are islands or collections of islands) are
considered neighbours to their *k* nearest provinces. For this example,
we have set the value of k to 2. The resulting neighbourhood structure
is shown in Figure \@ref(fig:`r ifelse(knitr::is_html_output(), 'ind-contig1html', 'ind-contig1')`).

(ref:my-caption1) Neighbourhood structure for Indonesian provinces created by `st_bridges()` with `k=2`.

```{r, ind-contig1, fig.width=12, fig.height=6, echo=TRUE, out.width="100%", eval=knitr::is_latex_output(), fig.cap = "(ref:my-caption1)"}
# join islands to k=2 nearest neighbours
# various arguments exist for altering colours and sizes
# additional ggplot themes and layers can be added

st_bridges(provinces_df, "province",link_islands_k = 2) |> 
  st_quickmap_nb(fillcol = "antiquewhite1", 
                 bordercol = "black", bordersize = 0.5, 
                 linkcol = "darkblue", linksize = 0.8, 
                 pointcol = "red", pointsize = 2) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank()) +
  geom_sf(data=nearby_countries_df, 
          fill="gray50", linewidth=0.5, colour="black")

```

```{r, ind-contig1html, fig.width=12, fig.height=6, echo=FALSE, out.width="100%", eval=knitr::is_html_output(), fig.cap = "Neighbourhood structure for Indonesian provinces created by st_bridges() with k=2. "}
# join islands to k=2 nearest neighbours
# various arguments exist for altering colours and sizes
# additional ggplot themes and layers can be added

st_bridges(provinces_df, "province",link_islands_k = 2) |> 
  st_quickmap_nb(fillcol = "antiquewhite1", 
                 bordercol = "black", bordersize = 0.5, 
                 linkcol = "darkblue", linksize = 0.8, 
                 pointcol = "red", pointsize = 2) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank()) +
  geom_sf(data=nearby_countries_df, 
          fill="gray50", linewidth=0.5, colour="black")

```

This neighbourhood structure now has no unconnected provinces so it is
suitable for use in an ICAR model. However, if we are not entirely happy with this structure because of some domain knowledge about the inter-relationships between certain island provinces, we might wish to

-   add some additional contiguities using `st_manual_join_nb()`

-   and remove one using `st_manual_cut_nb()`.

To cater for the possibility that a modeller might not be familiar with the names of the various geographic units but still wishes to make manual alterations to their relationships, we can look at a map (Figure \@ref(fig:`r ifelse(knitr::is_html_output(), 'ind-contig2html', 'ind-contig2')`)) where the nodes are shown by
index number instead of as points (using the argument `nodes='numeric'`). This makes it easy to manually cut
and join neighbours as desired. Furthermore, there is an option to
show concave hulls drawn around each unit (using `concavehull = TRUE`). This is also shown in Figure \@ref(fig:`r ifelse(knitr::is_html_output(), 'ind-contig2html', 'ind-contig2')`). These shapes are not used in the assignment of contiguities but it can be useful to see them in a situation such as Indonesia where many individual provinces are actually multipolygons of more than one island. Without them, it is not clear whether an island is a province in its own right, or which group of islands together form one province.

(ref:my-caption2) Neighbourhood structure for Indonesian provinces. Viewed with `st_quickmap_nb()`, using the arguments `nodes = 'numeric'` and `concavehull = TRUE`.

```{r, ind-contig2, fig.width=12, fig.height=6, echo=TRUE, out.width="100%", eval=knitr::is_latex_output(), fig.cap = "(ref:my-caption2)"}
# with 'concavehull = TRUE' and 'nodes = "numeric"'

st_bridges(provinces_df, "province",link_islands_k = 2) |> 
  st_quickmap_nb(fillcol = "antiquewhite1", 
                 bordercol = "black", bordersize = 0.5, 
                 linkcol = "tomato", linksize = 0.5, 
                 nodes = "numeric", 
                 numericcol = "black", numericsize = 6, 
                 concavehull = TRUE, 
                 hullcol = "darkgreen", hullsize = 0.2) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank())

```

```{r, ind-contig2html, fig.width=12, fig.height=6, echo=TRUE, out.width="100%", eval=knitr::is_html_output(), fig.cap = "Neighbourhood structure for Indonesian provinces. Viewed with st\\_quickmap_nb(), using the arguments nodes = 'numeric' and concavehull = TRUE. "}
# with 'concavehull = TRUE' and 'nodes = "numeric"'

st_bridges(provinces_df, "province",link_islands_k = 2) |> 
  st_quickmap_nb(fillcol = "antiquewhite1", 
                 bordercol = "black", bordersize = 0.5, 
                 linkcol = "tomato", linksize = 0.5, 
                 nodes = "numeric", 
                 numericcol = "black", numericsize = 6, 
                 concavehull = TRUE, 
                 hullcol = "darkgreen", hullsize = 0.2) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank())

```

Having made some manual adjustments to the neighbourhood structure, the
new structure can be seen in Figure \@ref(fig:`r ifelse(knitr::is_html_output(), 'ind-contig3html', 'ind-contig3')`). Edge effects
have also been mitigated by imposing additional connections on the two extreme
provinces (1 and 23), which would otherwise have only one neighbour, so that they now also include
their two next closest neighbours.

(ref:my-caption3) Neighbourhood structure for Indonesian provinces, after alterations using `st_manual_join()` and `st_manual_cut()`.

```{r, ind-contig3, fig.width=12, fig.height=6, echo=TRUE, out.width="100%", eval=knitr::is_latex_output(), fig.cap = "(ref:my-caption3)"}
# a series of manual joins and cuts by index number

st_bridges(provinces_df, "province",link_islands_k = 2) |> 
  st_manual_join_nb(1,24) |> 
  st_manual_join_nb(1,30) |> 
  st_manual_join_nb(3,13) |> 
  st_manual_join_nb(13,17) |> 
  st_manual_join_nb(14,25) |> 
  st_manual_join_nb(20,29) |>
  st_manual_join_nb(19,23) |> 
  st_manual_join_nb(16,27) |> 
  st_manual_join_nb(22,23) |> 
  st_manual_join_nb(7,19) |> 
  st_manual_join_nb(7,20) |>
  st_manual_join_nb(19,28) |> 
  st_manual_join_nb(4,18) |> 
  st_manual_join_nb(21,26) |> 
  st_manual_join_nb(22,28) |> 
  st_manual_cut_nb(19,22) |> 
  st_quickmap_nb(fillcol = "antiquewhite1", 
                 bordercol = "black", bordersize = 0.5, 
                 linkcol = "darkblue", linksize = 0.8, 
                 pointcol = "red", pointsize = 2) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank()) +
  geom_sf(data=nearby_countries_df, 
          fill="gray50", linewidth=0.5, colour="black") + 
  annotation_scale()

```

```{r, ind-contig3html, fig.width=12, fig.height=6, echo=FALSE, out.width="100%", eval=knitr::is_html_output(), fig.cap = "Neighbourhood structure for Indonesian provinces, after alterations using st\\_manual_join() and st\\_manual_cut(). "}
# a series of manual joins and cuts by index number

st_bridges(provinces_df, "province",link_islands_k = 2) |> 
  st_manual_join_nb(1,24) |> 
  st_manual_join_nb(1,30) |> 
  st_manual_join_nb(3,13) |> 
  st_manual_join_nb(13,17) |> 
  st_manual_join_nb(14,25) |> 
  st_manual_join_nb(20,29) |>
  st_manual_join_nb(19,23) |> 
  st_manual_join_nb(16,27) |> 
  st_manual_join_nb(22,23) |> 
  st_manual_join_nb(7,19) |> 
  st_manual_join_nb(7,20) |>
  st_manual_join_nb(19,28) |> 
  st_manual_join_nb(4,18) |> 
  st_manual_join_nb(21,26) |> 
  st_manual_join_nb(22,28) |> 
  st_manual_cut_nb(19,22) |> 
  st_quickmap_nb(fillcol = "antiquewhite1", 
                 bordercol = "black", bordersize = 0.5, 
                 linkcol = "darkblue", linksize = 0.8, 
                 pointcol = "red", pointsize = 2) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank()) +
  geom_sf(data=nearby_countries_df, 
          fill="gray50", linewidth=0.5, colour="black") + 
  annotation_scale()

```

## `mgcv` model

We now create the ICAR model using, in this case, the \CRANpkg{mgcv} package. We will be able to use the output of `st_bridges`, which we have named `prep_data`, as both the data source for the model and the neighbourhood structure (by specifying the column `nb` which contains the neighbourhood list).

```{r, fig.width=12, fig.height=6}

provinces_df$province <- factor(provinces_df$province)

prep_data <- st_bridges(provinces_df, "province",link_islands_k = 2) |> 
  st_manual_join_nb(1,24) |> 
  st_manual_join_nb(1,30) |> 
  st_manual_join_nb(3,13) |> 
  st_manual_join_nb(13,17) |> 
  st_manual_join_nb(14,25) |> 
  st_manual_join_nb(20,29) |>
  st_manual_join_nb(19,23) |> 
  st_manual_join_nb(16,27) |> 
  st_manual_join_nb(22,23) |> 
  st_manual_join_nb(7,19) |> 
  st_manual_join_nb(7,20) |>
  st_manual_join_nb(19,28) |> 
  st_manual_join_nb(4,18) |> 
  st_manual_join_nb(21,26) |> 
  st_manual_join_nb(22,28) |> 
  st_manual_cut_nb(19,22)

```

```{r, echo=TRUE}

mod_pois_mrf <- gam(damaging_quakes_total ~ 
                      fault_concentration +
                      s(province, bs='mrf', xt=list(nb=prep_data$nb), k=22) +
                      offset(log(area_province)),
                    data=prep_data, method="REML",family = "poisson")

```

We can see from the summary below that the adjusted R-squared is **0.957** and deviance explained is **88.2%**.

```{r}
summary(mod_pois_mrf)
```

Controlling for the concentration of faults, what is the additional risk level of
earthquakes in a province? This can be seen as a measure of the activity
level of faults locally and it is spatially smoothed by the
autoregressive process. It is represented in the model summary by the component `s(province)`. However, the extraction of individual predictions for this component for each province from the `mgcv` model requires a number of steps. We know demonstrate how these are streamlined into a single function by the `sfislands` package.

```{r, eval=FALSE}

# these two are equivalent
AER::dispersiontest(mod_pois_mrf)
# AER::dispersiontest(mod_pois_mrf,trafo = 1)

# these two are equivalent
# AER::dispersiontest(mod_pois_mrf,trafo = function(x) x^2)
# AER::dispersiontest(mod_pois_mrf,trafo = 2)

```

## Post-functions

The function `st_augment()` allow us to add the spatially varying predictions from the model as new columns to the original dataframe in a process
similar to that of the \CRANpkg{broom} package. For instance, we see from the code below that the original dataframe is now augmented with columns called "mrf.smooth.province" and "se.mrf.smooth.province" which show the predictions for the $\gamma_i$ component and their standard errors. Note that this is how we would expect them to be named, based on the previous discussion surrounding Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:staugtab-html)', '\\@ref(tab:staugtab-latex)'))`. They are positioned immediately before the final `geometry` column of the `sf` dataframe.

```{r, echo=TRUE}
# column names of augmented dataframe

mod_pois_mrf |> 
  st_augment(prep_data) |> 
  names()
```

This output can now be piped into the `st_quickmap_preds()` function to get a quick visualisation of
these estimates for $\gamma_i$ on a map, as shown in Figure \@ref(fig:`r ifelse(knitr::is_html_output(), 'ind-mrf1html', 'ind-mrf1')`). Again, note that the title and subtitle of the image are as previously discussed.

(ref:my-caption4) Estimates of $\gamma_i$ shown as a map using `st_quickmap_preds()`.

```{r, ind-mrf1, fig.width=12, fig.height=6, echo=TRUE, out.width="100%", eval=knitr::is_latex_output(), fig.cap = "(ref:my-caption4)"}
# st_quickmap_preds() outputs a list of ggplots

plot_mrf <- mod_pois_mrf |> 
  st_augment(prep_data) |>
  st_quickmap_preds(scale_low = "darkgreen",
                    scale_mid = "ivory", 
                    scale_high = "darkred", 
                    scale_midpoint = 0)

# in this case, there is only one plot in the list
# so we call it by index

plot_mrf[[1]] +
  coord_sf(datum=NA) +
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank()) +
  geom_sf(data=provinces_df, fill=NA, colour="black", linewidth=0.5) + 
  geom_sf(data=nearby_countries_df, fill="gray50", colour="black", 
          linewidth=0.5) + 
  labs(fill="relative\nincidence") +
  annotation_scale() +
  coord_sf(datum=NA) + 
  theme(legend.position = c(0.92,0.77),
        legend.box.background = element_rect(colour = "black", linewidth = 1),
        legend.title = element_text())

```

```{r, ind-mrf1html, fig.width=12, fig.height=6, echo=FALSE, out.width="100%", eval=knitr::is_html_output(), fig.cap = "Estimates of $\\gamma_i$ using st_quickmap_preds(). "}
# st_quickmap_preds() outputs a list of ggplots

plot_mrf <- mod_pois_mrf |> 
  st_augment(prep_data) |>
  st_quickmap_preds(scale_low = "darkgreen",
                    scale_mid = "ivory", 
                    scale_high = "darkred", 
                    scale_midpoint = 0)

# in this case, there is only one plot in the list

plot_mrf[[1]] +
  coord_sf(datum=NA) +
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank()) +
  geom_sf(data=provinces_df, fill=NA, colour="black", linewidth=0.5) + 
  geom_sf(data=nearby_countries_df, fill="gray50", colour="black", 
          linewidth=0.5) + 
  labs(fill="relative\nincidence") +
  annotation_scale() +
  coord_sf(datum=NA) + 
  theme(legend.position = c(0.92,0.77),
        legend.box.background = element_rect(colour = "black", linewidth = 1),
        legend.title = element_text())

```

If we wish to apply the inverse link function (the exponential function in the case of this poisson model) to map these values to a more interpretable scale, this will not be generated by the function `st_quickmap_preds()`. Instead, we must use the augmented dataframe which is produced by `st_augment()` and create the appropriate extra column with the usual \CRANpkg{tidyverse} `mutate()` function. This allows us to produce the map in Figure
\@ref(fig:`r ifelse(knitr::is_html_output(), 'ind-mrf2html', 'ind-mrf2')`). As these coefficients are multiplicatively related
to the earthquake incidence, values below 1 imply an earthquake
incidence which is lower than expected.

The provinces with the 3 most elevated incidences are labelled in red
We can see that, controlling for the effects of proximity to
faults, the province of Aceh has 7.9 times the expected
incidence, or number of major earthquakes per square kilometre. The two
lowest-scoring provinces, labelled in green, have essentially no incidence of
earthquake epicentres within their boundaries, controlling for what their proximity to faults alone would
suggest.

(ref:my-caption5) Map showing estimates of $\text{exp}(\gamma_i).$ This is produced by adding an additional column to the dataframe produced by `st_augment()`.

```{r, ind-mrf2, fig.width=12, fig.height=6, out.width="100%", eval=knitr::is_latex_output(), fig.cap = "(ref:my-caption5)"}

qpois_mod_aug <- mod_pois_mrf |> 
  st_augment(prep_data)

largest <- qpois_mod_aug |> 
  arrange(desc(mrf.smooth.province)) |> 
  head(3)
smallest <- qpois_mod_aug |> 
  arrange(mrf.smooth.province) |> 
  head(2)

ggplot() + 
  geom_sf(data=nearby_countries_df, 
          fill="gray50", colour="black", linewidth=0.5) + 
  geom_sf(data=qpois_mod_aug, aes(fill=exp(mrf.smooth.province)), 
          colour="black", linewidth=0.5) + 
  geom_sf_label_repel(data=largest, 
                      aes(label=paste(province,":",
                                      round(exp(mrf.smooth.province),1))),
                      colour="darkred", size=2.5,
                      nudge_y= 300000) +
  geom_sf_label_repel(data=smallest |> mutate(geometry=st_centroid(geometry)), 
               aes(label=paste(province,":",round(exp(mrf.smooth.province),3))),
               nudge_y= 200000, colour="darkgreen", size=2.5) +
  scale_fill_gradient2(low = "darkgreen", 
                       mid = "ivory", 
                       high = "darkred", 
                       midpoint = 1) +
  scale_colour_gradient2(low = "darkgreen", 
                         mid = "ivory", 
                         high = "darkred", 
                         midpoint = 1) +
  guides(colour="none") + 
  labs(fill="exp.\nrelative\nincidence") + 
  annotation_scale() +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", linewidth=1.5), 
        axis.title = element_blank(),
        axis.ticks = element_blank(), 
        legend.position = c(0.92,0.76),
        legend.box.background = element_rect(colour = "black", linewidth = 1))

```

```{r, ind-mrf2html, fig.width=12, fig.height=6, out.width="100%", eval=knitr::is_html_output(), fig.cap = "Map showing estimates of exp($\\gamma_i$). This is produced by adding an additional column to the dataframe produced by st_augment()."}

qpois_mod_aug <- mod_pois_mrf |> 
  st_augment(prep_data)

largest <- qpois_mod_aug |> 
  arrange(desc(mrf.smooth.province)) |> 
  head(3)
smallest <- qpois_mod_aug |> 
  arrange(mrf.smooth.province) |> 
  head(2)

ggplot() + 
  geom_sf(data=nearby_countries_df, 
          fill="gray50", colour="black", linewidth=0.5) + 
  geom_sf(data=qpois_mod_aug, aes(fill=exp(mrf.smooth.province)), 
          colour="black", linewidth=0.5) + 
  geom_sf_label_repel(data=largest, 
                      aes(label=paste(province,":",
                                      round(exp(mrf.smooth.province),1))),
                      colour="darkred", size=2.5,
                      nudge_y= 300000) +
  geom_sf_label_repel(data=smallest |> mutate(geometry=st_centroid(geometry)), 
               aes(label=paste(province,":",round(exp(mrf.smooth.province),3))),
               nudge_y= 200000, colour="darkgreen", size=2.5) +
  scale_fill_gradient2(low = "darkgreen", 
                       mid = "ivory", 
                       high = "darkred", 
                       midpoint = 1) +
  scale_colour_gradient2(low = "darkgreen", 
                         mid = "ivory", 
                         high = "darkred", 
                         midpoint = 1) +
  guides(colour="none") + 
  labs(fill="exp.\nrelative\nincidence") + 
  annotation_scale() +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#ECF6F7", colour = "black", linewidth=1.5), 
        axis.title = element_blank(),
        axis.ticks = element_blank(), 
        legend.position = c(0.92,0.76),
        legend.box.background = element_rect(colour = "black", linewidth = 1))

```

## Workflow summary

In this example, we have gone through a number of stages carefully, making changes to contiguities that we deemed appropriate as we went. However, in practice, at least in a first iteration, it might not be necessary to go through all of these steps. A rough and ready model, complete with spatially varying coefficients and visual output, can be generated with `sfislands` using nothing more than three or four lines of code, such as the following:

```{r, eval=FALSE, echo=TRUE}
# workflow:

# 1. set up neighbourhood structure

prep_data <- st_bridges(provinces_df, "province")

# 2. define model

mod <- gam(quake_mlxl_total ~ 
                  fault_concentration +
                  s(province, bs='mrf', xt=list(nb=prep_data$nb), k=22) +
                  offset(log(area_province)),
        data=prep_data, method="REML",family = "poisson")

# 3. augment tidy estimates

tidy_ests <- st_augment(mod, prep_data)

# 4. visualise them

st_quickmap_preds(tidy_ests)
```

# Example 2: London

The next example looks only at using the *pre-functions* of `sfislands`, but in a situation where the presence of actual *islands* is not the problem we seek to deal with. Consider the wards and boroughs of London (sourced from the Greater London Authority's [London Datastore](https://data.london.gov.uk/)). In Figure \@ref(fig:lon-contig1) the `st_bridges()`
function is applied to them to construct a queen contiguity
neighbourhood structure. Because there are no disjoint units (or
"islands"), this will be the same as using `st_contiguity()` from
\CRANpkg{sfdep}. The `st_quickmap_nb()` function gives an immediate
visual representation of the structure. ^[`st_quickmap_nb()` can also be used to visualise any contiguity structure created by `spdep` or `sfdep` as long as that structure is included in an `sf` dataframe as a column named `nb`.] This can be supplemented by
adding a layer showing the course of the river Thames which is also visible in Figure \@ref(fig:lon-contig1).

```{r, lon-contig1, fig.width=12, fig.height=8, echo=TRUE, out.width="100%", fig.cap = "Wards and boroughs of Greater London. Queen contiguity. "}
# same as sfdep:st_contiguity() as there are no islands
# an extra layer for the river Thames

st_bridges(london, "NAME") |> 
  st_quickmap_nb() + 
  geom_sf(data=thames, colour="blue", linewidth=1.5) + 
  theme(panel.background = element_rect(fill = "#F6F3E9", colour = "black", 
                                        linewidth=1.5))

```

When a study area has a river running through it, problems can arise with constructing appropriate neighbourhood structures. Depending on how the
geometries are defined, the presence of a river can cause problems in two ways. In one situation, the river could be expressed as a polygon in its own right meaning that it severs any potential contiguity between units on either side of its banks. In this situation, no spatial units will be neighbours with the units directly across the river from them. At the other extreme, if the river is not included as a geometry (as is the case here) all units on opposing banks are considered neighbours.

Depending on the presence of river crossings, two areas which are physically quite close but on opposing banks might be very distinct. If there is no means of crossing the river within a reasonable distance, somebody living on the banks of a river might be more likely to go about their life primarily on their side of the river, despite the short distance as the crow flies of facilities on the other side. This could be relevant in terms of, say, modelling of house prices where we might want to incorporate these issues into a neighbourhood structure.

`sfislands` provides convenient functions for this sort of situation. Let us start by restricting our wards and boroughs of interest to just those which are on either side of the river Thames (Figure \@ref(fig:lon-contig2)).

```{r, lon-contig2, fig.width=12, fig.height=6, echo=TRUE, out.width="100%", fig.cap = "Riverside boroughs of Greater London. Queen contiguity disregarding river. "}
# which boroughs are alongside the river

riverside <- thames |> st_intersects(london) |> unlist() |> unique()

# only map these boroughs

st_bridges(london[riverside,],"NAME") |> 
  st_quickmap_nb(linksize = 0.5) +
  geom_sf(data=thames, colour="blue", linewidth=1.5) + 
  annotation_scale(location="br") +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#F6F3E9", colour = "black", 
                                        linewidth=1.5))
```

In order to take account of actual connectivity, we can add a layer
showing the road and pedestrian bridges or tunnels. Details of these were sourced @wikicrossings, which describes the various crossings of the river Thames. In Figure \@ref(fig:lon-crossings), we have also drawn a 1 kilometre buffer around each crossing. This was chosen as an arbitrary measure of what might be considered a "reasonable" distance within which to consider opposing banks as being connected. The vast majority of units on opposing banks have access to a river crossing within this threshold and thus should be considered as neighbours. Only the extreme eastern units and one to the south west should not have a connection across the river.

```{r, lon-crossings, fig.width=12, fig.height=6, out.width="100%", fig.cap = "Riverside boroughs of Greater London. Road and pedestrian crossing and tunnels surrounded by 1 kilometre buffer shaded green. "}

ggplot() + 
  geom_sf(data=london[riverside,], fill="gray90", colour="gray50", linewidth=0.5) + 
  geom_sf(data=crossings_roadped |> st_buffer(1000), fill="darkgreen", alpha=0.4) + 
  geom_sf(data=thames, colour="blue", linewidth=1.5) + 
  geom_sf(data=crossings_roadped, size=1, colour="yellow") + 
  geom_sf_label_repel(data=crossings_roadped[1:11,], aes(label=crossing), size=2.5, nudge_y=-10000, fill="white", fontface="bold") + 
  geom_sf_label_repel(data=crossings_roadped[12:22,], aes(label=crossing), size=2.5, nudge_y=10000, fill="white", fontface="bold") +
  annotation_scale(location="br") +
  coord_sf(datum=NA)

```

```{r}
# which riverside units do not fall within any of the buffers...

crossing_buffer_summarised <- crossings_roadped |> 
  st_buffer(1000) |> 
  st_transform(st_crs(london)) |> 
  summarise()

touch_buffer <- st_intersects(london[riverside,],
                              crossing_buffer_summarised)
indices <- c()
for (i in seq_along(touch_buffer)) {
  if (length(touch_buffer[[i]]) == 0) {
    indices <- c(indices, i)
  }
}
indices <- which(sapply(touch_buffer, length) == 0)

no_touch_buffer <- london[riverside,][indices,]
```

In order the identify the changes we wish to make, we use the
`nodes = "numeric"` argument in `st_quickmap_nb()`. Now we can identify
each unit by its position in the contiguity structure. Here we have
shaded in pink the units which are not within 1 kilometre of a river
crossing (see Figure \@ref(fig:lon-contig3)).

```{r, lon-contig3, fig.width=12, fig.height=6, echo=TRUE, out.width="100%", fig.cap = "Riverside boroughs of Greater London. Index number for each borough shown at centroid. Boroughs which are not within 1 kilometre of a crossing are shaded pink. "}
# with 'nodes = "numeric"'

st_bridges(london[riverside,],"NAME") |> 
  st_quickmap_nb(nodes = "numeric", numericsize = 4, linksize = 0.5) +
  geom_sf(data=no_touch_buffer, fill="pink", alpha=0.3) + 
  geom_sf(data=crossings_roadped |> st_buffer(1000), 
          fill="darkgreen", alpha=0.3) +
  geom_sf(data=thames, colour="blue", linewidth=1.5) + 
  geom_sf(data=crossings_roadped, size=1, colour="yellow") + 
  annotation_scale(location="br") +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#F6F3E9", colour = "black", 
                                        linewidth=1.5))
```

This allows us to easily cut the ties across the river for these units
manually by using the function `st_manual_cut_nb()`. ^[While we are using the index of the units in this example, the function also
accepts names as arguments which may be more convenient in some circumstances.] Having made these adjustments,
`st_quickmap_nb()` now shows a connectivity structure (Figure \@ref(fig:lon-contig4)) which reflects how we have hypothesised that influence should extend across the river in the presence or absence of crossings.

This example shows that the pre-functions of `sfislands` have uses for situations which do not involve islands. They can be used to apply domain knowledge to easily design the most appropriate neighbourhood structure.

```{r, lon-contig4, fig.width=12, fig.height=6, echo=TRUE, out.width="100%", fig.cap = "Riverside boroughs of Greater London. Contiguities across the river have been cut for the pink boroughs. "}
# manually cut the links where there is no crossing

st_bridges(london[riverside,], "NAME") |> 
  st_manual_cut_nb(18,17) |> 
  st_manual_cut_nb(19,17) |> 
  st_manual_cut_nb(19,20) |> 
  st_manual_cut_nb(20,21) |> 
  st_manual_cut_nb(21,22) |> 
  st_manual_cut_nb(22,48) |> 
  st_manual_cut_nb(47,48) |> 
  st_manual_cut_nb(45,46) |> 
  st_manual_cut_nb(45,47) |>  
  st_manual_cut_nb(39,65) |> 
  st_manual_cut_nb(1,2) |> 
  st_quickmap_nb(bordercol = "black", bordersize = 0.5, linksize = 0.5) +
  geom_sf(data=no_touch_buffer, fill = "pink", alpha = 0.3) + 
  geom_sf(data=crossings_roadped |> st_buffer(1000), 
          fill= "darkgreen", alpha = 0.3) +
  geom_sf(data=thames, colour = "blue", linewidth = 1.5) +
  annotation_scale(location = "br") +
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#F6F3E9", colour = "black", 
                                        linewidth = 1.5))
```

# Example 3: Liverpool

In this final example, there are no islands and no river issues. We saw in the Indonesia example how `sfislands` can extract estimates from the `mgcv` argument `bs="mrf"`. It can also deal with random effects as defined by `bs="re"` in `mgcv` models. This means it is possible to create quite complex models featuring both a traditional multi-level structure and conditionally related components. Here, we
will construct a hierarchical model in this way which also features an ICAR component at the lowest level of the tree. We will
compare two different types of contiguity which could be applied to the ICAR component:

* `st_bridges()` from `sfislands`, which implements first-order queen contiguity in the same way as `sfdep:st_contiguity()` for situations such as this where there are no islands, and 
* `st_dist_band()` from \CRANpkg{sfdep}, which connects units whose centroids lie within a chosen distance threshold.

We will show how their outcomes can be quickly mapped to give an immediate sense of the differences implied by each model's structure.

This example is based on data from the [Office of National Statistics](https://www.nomisweb.co.uk/home/census2001.asp) as presented by @liverpool-notes in their demonstration of techniques for hierarchical spatial regression. They use the different levels of nested administrative boundaries in Liverpool for their demonstration. These are described in Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:livadmin-html)', '\\@ref(tab:livadmin-latex)'))`, and are mapped separately in Figure \@ref(fig:liv-nested).

Now suppose we wish to model unemployment as a function of limiting long-term
illness, taking into account this hierarchical structure. Limiting long-term illness (LLTI) refers to a health problem or disability which limits a person's day-to-day activities [@Barnett2001]. We wish to incorporate a fixed intercept and coefficient for LLTI, with additional random intercepts and slopes at MSOA and LSOA levels. The raw distribution of these variables across the output areas (OAs) of Liverpool are shown in Figure
\@ref(fig:liv-vars).

If $y_{ijk}$ and $x_{ijk}$ are respectively the levels of unemployment and LLTI in each of $i$ OAs within $j$ lower level super output areas (LSOAs), which in turn are within $k$ middle level super output areas (MSOAs), then 
\begin{equation}
y_{ijk} = \beta_{0jk} + \beta_{1jk}x_{ijk} + \gamma_{ijk} + \epsilon_{ijk}
(\#eq:eq9)
\end{equation}
\begin{equation}
\beta_{0jk} = \beta_{0k} + u_{0jk}
(\#eq:eq10)
\end{equation}
\begin{equation}
\beta_{1jk} = \beta_{1k} + u_{1jk}
(\#eq:eq11)
\end{equation}
\begin{equation}
\beta_{0k} = \beta_0 + w_{0k}
(\#eq:eq12)
\end{equation}
\begin{equation}
\beta_{1k} = \beta_1 + w_{1k}
(\#eq:eq13)
\end{equation}
\begin{equation}
\epsilon_{ijk} \sim N(0,\sigma_{\epsilon}^2)
(\#eq:eq14)
\end{equation}
\begin{equation}
u_{jk} \sim N(0,\sigma_{u}^2)
(\#eq:eq15)
\end{equation}
\begin{equation}
 w_{k} \sim N(0,\sigma_{w}^2)
(\#eq:eq16)
\end{equation}
where the $\gamma_{ijk}$s comprise a vector of ICAR components for each OA, similar in form to the ICAR component for the provinces in the earlier Indonesia model. The vector has a multivariate normal distribution with mean 0 and precision related to the contiguity structure as before.

There are no islands or discontiguities in this geometry of Liverpool so
using `st_bridges()` will give the same structure as would
`sfdep::st_contiguity()` (see Figure \@ref(fig:liv-contigs)).

```{r}
livadmin <- tribble(
  ~code, ~`area type`, ~description,
  "OA", "output areas", "lowest level of geographical area for census statistics, usually containing 100 - 625 persons",
  "LSOA", "lower layer super output areas", "usually 4 or 5 OAs",
  "MSOA", "middle layer super output areas", "usually 4 or 5 LSOAs"
)

```

```{r livadmin-html, eval = knitr::is_html_output(), layout = "center"}
knitr::kable(livadmin, format = "html", caption = "Three nested levels of administrative divisions in Liverpool.") |> 
  kable_styling(full_width = TRUE)
```

```{r livadmin-latex, eval = knitr::is_latex_output()}
knitr::kable(livadmin, format = "latex", caption = "Three nested levels of administrative divisions in Liverpool.") |> 
  row_spec(0, bold = TRUE) |> 
  kableExtra::kable_styling(font_size = 8) |> 
  column_spec(3, width = "7cm")
```

```{r, liv-nested, fig.width=10, fig.height=5, out.width="100%", fig.cap = "Nested levels of administrative divisions in Liverpool. "}

ggarrange(
  ggplot() + 
    geom_sf(data=msoa_sf, fill="gray90", colour="black") +  
    geom_sf(data=mersey, fill="lightblue", colour="black", linewidth=0.2) + 
    geom_sf(data=liverpool |> summarise(), colour="black", fill=NA) + 
    labs(title = "level 1: MSOA polygons") + 
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#F7FFF2", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank()),

  ggplot() + 
    geom_sf(data=lsoa_sf, fill="gray90", colour="gray40") +  
    geom_sf(data=mersey, fill="lightblue", colour="black", linewidth=0.2) + 
    geom_sf(data=liverpool |> summarise(), colour="black", fill=NA) + 
    labs(title = "level 2: LSOA polygons") + 
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#F7FFF2", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank()),

  ggplot() + 
    geom_sf(data=liverpool |> summarise(), colour="black", fill="white") + 
    geom_sf(data=liverpool |> st_centroid(), 
            fill="gray90", colour="gray50", size=0.1) +  
    geom_sf(data=mersey, fill="lightblue", colour="black", linewidth=0.2) + 
    labs(title = "level 3: OA (shown as centroids)") + 
  coord_sf(datum=NA) + 
  theme(panel.background = element_rect(fill = "#F7FFF2", colour = "black", 
                                        linewidth=1.5),
        axis.text = element_blank()),

  ncol=3
)

```

```{r, liv-vars, fig.width=6, fig.height=4, out.width="70%", fig.cap = "Unemployment and limiting long-term illness rates in Liverpool, with extent of Mersey estuary in light blue to the south west. "}

ggarrange(
  ggplot() + 
    geom_sf(data=liverpool, aes(fill=unemployment, colour=unemployment)) + 
    geom_sf(data=liverpool |> summarise(), 
            fill=NA, colour="gray30", linewidth=0.2) +  
    geom_sf(data=mersey, fill="lightblue", colour="gray30", linewidth=0.2) + 
    scale_fill_distiller(palette = "PuBu", direction = 1) + 
    scale_colour_distiller(palette = "PuBu", direction = 1) + 
    labs(fill="unemployment",
         colour="unemployment") + 
    coord_sf(datum=NA) + 
    annotation_scale(location = "bl", width_hint = 0.4) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    theme(panel.background = element_rect(fill = "#F7FFF2", colour = "black", 
                                          linewidth=0.6), 
          axis.title = element_blank(),
          axis.ticks = element_blank(),
          legend.text = element_text(size=8),
          legend.title = element_text(size=8)),
  ggplot() + 
    geom_sf(data=liverpool, aes(fill=lt_illness, colour=lt_illness)) + 
    geom_sf(data=liverpool |> summarise(), 
            fill=NA, colour="gray30", linewidth=0.2) +  
    geom_sf(data=mersey, fill="lightblue", colour="gray30", linewidth=0.2) + 
    scale_fill_distiller(palette = "RdYlBu", direction = -1) + 
    scale_colour_distiller(palette = "RdYlBu", direction = -1) + 
    labs(fill="limiting long-term\nillness\n(z-score)",
         colour="limiting long-term\nillness\n(z-score)") + 
    coord_sf(datum=NA) +
    annotation_scale(location = "bl", width_hint = 0.4) +
    annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.2, "in"),
        style = north_arrow_fancy_orienteering) +
    theme(panel.background = element_rect(fill = "#F7FFF2", colour = "black", 
                                          linewidth=0.6), 
          axis.title = element_blank(),
          axis.ticks = element_blank(),
          legend.text = element_text(size=8),
          legend.title = element_text(size=8)),

  ncol=2, legend = "bottom"
)

```

```{r, liv-contigs, fig.width=5, fig.height=5, out.width="60%", echo=TRUE, fig.cap = "First order queen contiguity of output areas (OAs) in Liverpool with magnified inset. "}
# show st_bridges() connection with a magnified section

from <- c(xmin = -2.97, xmax = -2.95, ymin = 53.38, ymax = 53.39)
to <- c(-3.01,-2.96,53.32,53.35)

st_bridges(liverpool, "oa_cd") |> 
  st_quickmap_nb(pointsize = 0.1,
                 linksize = 0.1,
                 fillcol = "gray90") +  
  geom_sf(data=mersey, fill="lightblue", colour="gray30", 
          linewidth=0.2, alpha=0.5) + 
  geom_magnify(data=liverpool, 
               from = from, to = to, 
               corners = 0.1, shadow = TRUE, linewidth = 0.6) + 
  theme(panel.background = element_rect(fill = "#F7FFF2",
                                        colour = "black", linewidth=1))

```

To demonstrate how `sfislands` can be used to quickly compare the output of models, we show the output of three similar models. The first of these is a hierarchical model as described above, but with no ICAR component at the lowest level. Its spatially varying coefficients are shown in Figure \@ref(fig:liv-est1). Secondly, we fit the model described above using `st_bridges()` to generate first-order queen contiguities as the structure for the neighbourhood relationships. The coefficients of this model are shown in Figure \@ref(fig:`r ifelse(knitr::is_html_output(), 'liv-est2html', 'liv-est2')`).

Finally, we can fit a similar model but this time using a within-distance criterion of 700m for neighbourhood status. All units within this distance of each other are considered neighbours. This was generated with the `sfdep:st_dist_band()` function. These coefficients are shown in Figure \@ref(fig:`r ifelse(knitr::is_html_output(), 'liv-est3html', 'liv-est3')`).

This shows how `sfislands` can quickly give an intuitive representation of how a set of similar models are operating. As a measure of the relative strengths of these three structures, we could compare their AICs (see Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:aic-html)', '\\@ref(tab:aic-latex)'))`). These suggest that, in this scenario, inclusion of the ICAR component is appropriate, and queen contiguity is a more suitable choice than the distance-band condition for defining the neighbourhood structure.

```{r}

# these models take a while to run
# so i am using a saved version of their output
# the files are large so are not on github
# uncomment the model and run it to reproduce

prepliv <- st_bridges(liverpool, "oa_cd")

# liverpool_model2 <- gam(unemployment ~ lt_illness + 
#                          s(msoa_cd, bs="re") + 
#                          s(msoa_cd, lt_illness, bs="re") +
#                          s(lsoa_cd, bs="re") + 
#                          s(lsoa_cd, lt_illness, bs="re"), # ICAR constituency ICAR varying coefficients
#                        data=prepliv, method="REML")

# saveRDS(liverpool_model2, here("data","liverpool_model2.rds"))

liverpool_hmodel <- readRDS(here("data","liverpool_model2.rds"))

# summary(liverpool_hmodel)

```

```{r}

# these models take a while to run
# so i am using a saved version of their output
# the files are large so are not on github
# uncomment the model and run it to reproduce

# liverpool_hsamodel <- gam(unemployment ~ lt_illness +
#                          s(msoa_cd, bs="re") +
#                          s(msoa_cd, lt_illness, bs="re") +
#                          s(lsoa_cd, bs="re") +
#                          s(lsoa_cd, lt_illness, bs="re") +
#                          s(oa_cd, bs='mrf',
#                            xt=list(nb=prepliv$nb),k=20), # ICAR constituency ICAR varying coefficients
#                        data=prepliv, method="REML")

# saveRDS(liverpool_hsamodel, here("data","liverpool_hsamodel.rds")

liverpool_hsamodel <- readRDS(here("data","liverpool_hsamodel.rds"))

# summary(liverpool_hsamodel)

```

```{r}

liverpool_hmodel_aug <- liverpool_hmodel |> 
  st_augment(prepliv) # pipe into function to get estimates

liverpool_hsamodel_aug <- liverpool_hsamodel |> 
  st_augment(prepliv) # pipe into function to get estimates

```

```{r}

liverpool_hmodel_plots <- liverpool_hmodel |> 
  st_augment(prepliv) |> # pipe into function to get estimates
  st_quickmap_preds(borderwidth = NA, 
                    scale_low = "darkblue",
                    scale_mid = "gray75",
                    scale_high = "darkred",
                    scale_midpoint = 0, 
                    framefill = "#F7FFF2",
                    frameline = "black",
                    framesize = 1) # pipe into this for visualisation

```

```{r, liv-est1, fig.width=12, fig.height=3, out.width="100%", fig.cap = "Random intercept and slope estimates from hierarchical model. Middle and lower layer super output areas as levels. "}
# put an empty plot on the right to line up with those with an extra lowest level

ggarrange(
  plotlist = liverpool_hmodel_plots,
  nrow = 1, legend = "right"
)

```

```{r}

liverpool_hsamodel_plots <- liverpool_hsamodel |> 
  st_augment(prepliv) |> # pipe into function to get estimates
  st_quickmap_preds(borderwidth = NA, 
                    # legendlimits = "minmax",
                    scale_low = "darkblue",
                    scale_mid = "gray75",
                    scale_high = "darkred",
                    scale_midpoint = 0, 
                    framefill = "#F7FFF2",
                    frameline = "black",
                    framesize = 1) # pipe into this for visualisation

```

(ref:my-caption6) Random intercept and slope estimates from hierarchical model. Middle and lower layer super output areas are the top two levels respectively (MSOA & LSOA). Output areas (OA) as lowest level Markov random field smooth using first-degree queen contiguity from `st_bridges()`.

```{r, liv-est2, fig.width=12, fig.height=6, out.width="100%", eval=knitr::is_latex_output(), fig.cap = "(ref:my-caption6)"}

ggarrange(
  ggarrange(
    liverpool_hsamodel_plots[[1]],
    liverpool_hsamodel_plots[[2]],
    liverpool_hsamodel_plots[[3]],
    liverpool_hsamodel_plots[[4]],
    nrow = 1
  ),
  liverpool_hsamodel_plots[[5]],
  nrow = 2
)

```

```{r, liv-est2html, fig.width=12, fig.height=6, out.width="100%", eval=knitr::is_html_output(), fig.cap = "Random intercept and slope estimates from hierarchical model. Middle and lower layer super output areas are the top two levels respectively (MSOA & LSOA). Output areas (OA) as lowest level Markov random field smooth using first-degree queen contiguity from st\\_bridges(). "}

ggarrange(
  ggarrange(
    liverpool_hsamodel_plots[[1]],
    liverpool_hsamodel_plots[[2]],
    liverpool_hsamodel_plots[[3]],
    liverpool_hsamodel_plots[[4]],
    nrow = 1
  ),
  liverpool_hsamodel_plots[[5]],
  nrow = 2
)

```

```{r}

# 0.7km distance band
prepliv <- liverpool |> 
  mutate(nb = st_centroid(liverpool) |> 
              st_dist_band(upper = 0.7))
names(prepliv$nb) <- prepliv$oa_cd

# these models take a while to run
# so i am using a saved version of their output
# the files are large so are not on github
# uncomment the model and run it to reproduce

# liverpool_hsamodel_dist_band <- gam(unemployment ~ lt_illness +
#                          s(msoa_cd, bs="re") +
#                          s(msoa_cd, lt_illness, bs="re") +
#                          s(lsoa_cd, bs="re") +
#                          s(lsoa_cd, lt_illness, bs="re") +
#                          s(oa_cd, bs='mrf',
#                            xt=list(nb=prepliv$nb),k=20), # ICAR constituency ICAR varying coefficients
#                        data=prepliv, method="REML")

# saveRDS(liverpool_hsamodel_dist_band, here("data","liverpool_hsamodel_dist_band.rds"))

liverpool_hsamodel_dist_band <- readRDS(here("data","liverpool_hsamodel_dist_band.rds"))

# summary(liverpool_hsamodel_dist_band)

```

```{r}

liverpool_hsamodel_dist_band_plots <- liverpool_hsamodel_dist_band |> 
  st_augment(prepliv) |> # pipe into function to get estimates
  st_quickmap_preds(borderwidth = NA, 
                    scale_low = "darkblue",
                    scale_mid = "gray75",
                    scale_high = "darkred",
                    scale_midpoint = 0, 
                    framefill = "#F7FFF2",
                    frameline = "black",
                    framesize = 1) # pipe into this for visualisation

```

(ref:my-caption7) Random intercept and slope estimates from hierarchical model. Middle and lower layer super output areas are the top two levels respectively (MSOA & LSOA). Output areas (OA) as lowest level Markov random field smooth using `sfdep::st_dist_band()` with radius of 700 metres.

```{r, liv-est3, fig.width=12, fig.height=6, out.width="100%", eval=knitr::is_latex_output(), fig.cap = "(ref:my-caption7)"}

ggarrange(
  ggarrange(
    liverpool_hsamodel_dist_band_plots[[1]],
    liverpool_hsamodel_dist_band_plots[[2]],
    liverpool_hsamodel_dist_band_plots[[3]],
    liverpool_hsamodel_dist_band_plots[[4]],
    nrow = 1
  ),
  liverpool_hsamodel_dist_band_plots[[5]],
  nrow = 2
)

```

```{r, liv-est3html, fig.width=12, fig.height=6, out.width="100%", eval=knitr::is_html_output(), fig.cap = "Random intercept and slope estimates from hierarchical model. Middle and lower layer super output areas are the top two levels respectively (MSOA & LSOA). Output areas (OA) as lowest level Markov random field smooth using sfdep\\:\\:st\\_dist_band() with radius of 700 metres."}

ggarrange(
  ggarrange(
    liverpool_hsamodel_dist_band_plots[[1]],
    liverpool_hsamodel_dist_band_plots[[2]],
    liverpool_hsamodel_dist_band_plots[[3]],
    liverpool_hsamodel_dist_band_plots[[4]],
    nrow = 1
  ),
  liverpool_hsamodel_dist_band_plots[[5]],
  nrow = 2
)

```

```{r}

aic_tab <- AIC(liverpool_hmodel, liverpool_hsamodel, liverpool_hsamodel_dist_band) |> 
  data.frame() |> 
  mutate(`relative AIC` = round(-AIC / min(AIC),3),
         model = c("hierachical model, no ICAR",
                   "hierachical model, ICAR based on st_bridges()",
                   "hierachical model, ICAR based on sfdep::st_dist_band()")) |> 
  rownames_to_column(var="modcode") |> 
  arrange(AIC) |> 
  select(model, AIC, `relative AIC`)

```

```{r aic-html, eval = knitr::is_html_output(), layout = "center"}
knitr::kable(aic_tab, format = "html", caption = "Comparison of models by AIC.") |> 
  kable_styling(full_width = TRUE)
```

```{r aic-latex, eval = knitr::is_latex_output()}
knitr::kable(aic_tab, format = "latex", caption = "Comparison of models by AIC.") |> 
  row_spec(0, bold = TRUE) |> 
  kableExtra::kable_styling(font_size = 8)
```

# Summary

These examples have shown the varying scenarios in which `sfislands` can be useful. It aims to contribute to spatial modelling by making an awkward area less awkward. Rather than having a default attitude of ignoring islands when building neighbourhood structures based on contiguity, it hopes to encourage at least an examination of whether or not it is appropriate for them to be included. It also provides helper functions to use these structures in spatial regression models built with `mgcv` which streamline the human effort necessary to examine the estimates.

# Acknowledgements

This publication has emanated from research conducted with the financial support of Science Foundation Ireland under Grant number 18/CRT/6049.