### 2024-04-16 README.Rmd
add link to arxiv preprint

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<img src="figures/logo.png" align="right" height="240"/>

<!-- badges: start -->
<!-- badges: end -->

This repository contains the code and data for the paper:

**sfislands: An R Package for Accommodating Islands and Disjoint Zones in Areal Spatial Modelling**.

A pre-print of the paper can be found on **arxiv** [here](https://arxiv.org/abs/2404.09863)